A faster exact method for solving the robust multi-mode resource-constrained project scheduling problem

### 2022-03-10 paper/paper.tex
Updates to paper.

% bibliography and referencing
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkgreen}{RGB}{0,127,173}
\definecolor{darkblue}{RGB}{0,0,200}
\hypersetup{colorlinks, linkcolor={darkblue}, citecolor={darkgreen}, urlcolor={darkblue}}

@@

\author[1]{Matthew Bold\footnote{Corresponding author, email: m.bold1@lancaster.ac.uk}}
\author[2]{Marc Goerigk}

\affil[1]{STOR-i Centre for Doctoral Training, Lancaster University, Lancaster, UK}	
\affil[2]{Network and Data Science Management, University of Siegen, Siegen, Germany}

@@

\begin{abstract}
\end{abstract}

@@

We propose solving this problem using an extended version of the compact formulation developed by \cite{bold2021compact} for solving the robust RCPSP. This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decompostion-based methods for the robust RCPSP.

This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network. This augmented project network is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project. Having reformulated the adversarial subproblem as a longest-path problem, it can be dualised an inserted into a formulation for the full problem, to arrive at a complete compact formulation for the first-stage problem. Given that the derivation of this formulation for the MRCPSP is more or less identical to the derivation for the RCPSP, we omit it here and refer the reader to \cite{bold2021compact}.

The resulting compact formulation for the two-stage adjustable robust MRCPSP can be written as
\begin{align}
\min \ & S_{n+1,\Gamma}\label{eqn:compact1}\\
\text{s.t. } & S_{00}=0\label{eqn:compact2}\\
	     & S_{j\gamma} - S_{i\gamma} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i,\,\gamma=0,\dots,\Gamma\label{eqn:compact_bigM1}\\
	     & S_{j,\gamma+1} - S_{i\gamma} \geq (\bar{d}_{im} + \hat{d}_{im})x_{im} - N(1-y_{ij})\nonumber \\
	     & & \hspace{-8cm} \forall (i,j)\in V^2,\, m\in M_i,\,\gamma = 0,\dots,\Gamma-1\label{eqn:compact_bigM2}\\
	     & y_{ij}=1 \qquad \forall (i,j) \in E\cup\{(n+1,n+1)\}\label{eqn:compact5}\\
	     & y_{ij} + y_{ji} \leq 1 \qquad \forall i,j\in V,\, i<j\label{eqn:compact_trans1}\\
	     & y_{ij} = y_{ip} + y_{p j} - 1 \qquad \forall i,j,p\in V,\,i\neq j\neq p\label{eqn:compact_trans2}\\
	     & f_{ijk}\leq P_{ijk}y_{ij} \qquad \forall (i,j)\in V^2,\,i\neq n+1,\,j\neq0,\, \forall k \in K\label{eqn:compact6}\\
	     & \sum_{i\in V\setminus\{n+1\}}f_{ijk}=\sum_{m\in M_j}r_{jmk}x_{jm} \qquad \forall j \in V\setminus\{0\} ,\, \forall k\in K\label{eqn:compact7}\\
	     & \sum_{j\in V\setminus\{0\}}f_{ijk}=\sum_{m\in M_i}r_{imk}x_{im} \qquad \forall i \in V\setminus\{n+1\},\,\forall k\in K\label{eqn:compact8}\\
	     & \sum_{m\in M_i}x_{im} = 1 \qquad \forall i\in V\label{eqn:compact9}\\
	     & \sum_{m\in M_i}r^{'}_{imk}x_{im} \leq R^{'}_k \qquad \forall i \in V,\,k\in K^{'}\label{eqn:compact_nonrenewable}\\
	     & S_{i\gamma}\geq 0 \qquad \forall i\in V,\,\gamma \in 0,\dots, \Gamma\label{eqn:compact10}\\
	     & f_{ijk}\geq 0 \qquad \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact11}\\
	     & y_{ij}\in\{0,1\} \qquad \forall (i,j)\in V^2\label{eqn:compact12}\\
	     & x_{im}\in\{0,1\} \qquad \forall i\in V,\,m\in M_i\label{eqn:compact13},
\end{align}
where $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan, and $P_{ijk}=\min\{\max_{m\in M_i}r_{imk},\,\max_{m\in M_j} r_{jmk}\}$ is the maximum possible flow of resource $k$ from $i$ into $j$.

@@

With polynomially many constraints this formulation can be implemented straightforwardly using standard mathematical optimisation software such as Gurobi or CPLEX.

@@

In this section we outline a Benders' decomposition approach to solving the robust MRCPSP first presented in \cite{balouka2021robust}. As mentioned previously, this approach is based on the approach used to solve the robust RCPSP in \cite{bruni2017adjustable}.

The main idea behind this approach is to decompose the full problem into a two smaller problems: 1. a master problem that determines activity processing modes and resolves resource conflicts, and 2. a subproblem that takes the solution from the master problem and evaluates it by finding the worst-case makespan for the resulting network by solving a longest path problem. The solution to the master problem forms a lower bound to the optimal objective value of the original problem, whilst the solution to the subproblem provides an upper bound. These two problems are solved iteratively, with optimality cuts being added to the master problem each iteration based on the solution to the subproblem until the lower and upper bounds meet .

@@

The role of the master problem is to determine a choice of activity processing modes $\bm{m}$, as well as a sufficient selection $X_{\bm{m}}$ to resolve any resulting resource conflicts. The processing mode for each activity is determined by binary variables $x_{im}$. The $y_{ij}$ variables define a complete set of precedences between the project activities, either due to the original project precedence constraints, or due to the resolution of resource conflicts, i.e. $\{(i,j),\, i,j\in V:y_{ij}=1\}=T(E\cup X_{\bm{m}})$, where $T(E\cup X_{\bm{m}})$ is the transitive closure of the extended project precedence network. Resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion. Finally, $s_i$ and $\phi_{ij}$ are additional variables used to define two relaxations of the subproblem to strengthen the master problem formulation. Using these variables, the master problem can be formulated as follows:  
\begin{align}
\min \ & \eta\label{eqn:master_first}\\
\text{s.t. } & \eta \geq S_{n+1}\\
	     & S_{0}=0\label{eqn:master2}\\
	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
	     & S_i\geq 0 \qquad \forall i\in V\label{eqn:master_last}
\end{align}

\begin{comment}
\begin{align}
	& \min \eta\label{eqn:master_first}\\
	& y_{ij} = 1 \qquad \forall (i,j)\in E \label{eqn:master_y1}\\
	& y_{ij} + y_{ji} \leq 1 \qquad \forall i,j\in V,\, i<j\label{eqn:master_y2}\\
	& y_{ij} = y_{ip} + y_{p j} - 1 \qquad \forall i,j,p\in V,\,i\neq j\neq p\label{eqn:master_y3}\\
	& f_{ijk} \leq \min(r_{im_ik}, r_{jm_jk})\cdot y_{ij} + (1 - x_{i,m_i})(\bar{r}_{ijk} - \min(r_{im_ik}, r_{jm_jk}))\nonumber\\
	& & \hspace{-8cm} + (1 - x_{j,m_j})(\bar{r}_{ijk} - \min(r_{im_ik}, r_{jm_jk})) \nonumber \\
	& & \hspace{-10cm}\forall i\in V\setminus\{n+1\},\, j \in V\setminus\{0\},\, i\neq j,\,m_i\in M_i,\,m_j\in M_j,\,k\in K\label{eqn:master_f1}\\
	& \sum_{i\in V\setminus\{j,n+1\}}f_{ijk} = \sum_{m\in M_j}r_{jmk}x_{jm} \qquad \forall j \in V\setminus\{0\} ,\,k\in K\label{eqn:master_f2}\\
	& \sum_{j\in V\setminus\{0,i\}}f_{ijk} = \sum_{m\in M_i}r_{imk}x_{im} \qquad \forall i \in V\setminus\{n+1\} ,\,k\in K\label{eqn:master_f3}\\
	& \sum_{m\in M_i}x_{im} = 1\qquad \forall i \in V\label{eqn:master_x}\\
	& \sum_{m\in M_i}r^{\nu}_{imk}x_{im} \leq R^{\nu}_k \qquad \forall i\in V,\,k\in K^{\nu}\label{eqn:master_nonrenewable}\\
	& \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=0,\dots,t-1\label{eqn:master_cuts}\\
	& s_j - s_i \geq \bar{d}_{im} - N(1-y_{ij}) - N(1-x_{im}) \qquad \forall i,j\in V,\, i\neq j,\,m\in M_i\label{eqn:master_relaxation1_1}\\
	& \eta \geq s_{n+1}\label{eqn:master_relaxation1_2}\\
	& \eta \geq \sum_{i \in V}\sum_{j\in V}\Big(\min_{m\in M_i}\bar{d}_{im}\Big)\phi_{ij}\label{eqn:master_relaxation2_1}\\
	& \sum_{j\in V}\phi_{ji} = \sum_{j\in V}\phi_{ij} \qquad \forall i \in V\setminus\{0,n+1\}\\
	& \sum_{j\in V}\phi_{0j} = 1\\
	& \sum_{j\in V}\phi_{j,n+1} = 1\\
	& \phi_{ij} \leq y_{ij} \qquad \forall i,j\in V,\,i\neq j\label{eqn:master_relaxation2_5}\\
	& y_{ij} \in\{0,1\} \qquad \forall i,j\in V,\,i\neq j\\
	& x_{im} \in \{0,1\} \qquad \forall i \in V,\, m\in M_i\\
	& f_{ijk} \geq 0 \qquad \forall i\in V\setminus\{n+1\},\,j\in V\setminus\{0\},\,i\neq j,\,k\in K\\
	& s_i \geq 0 \qquad \forall i\in V\\
	& \phi_{ij} \geq 0 \qquad \forall i,j\in V \label{eqn:master_last}
\end{align}
where $\bar{r}_{ijk}=\max(\max_{m\in M_i}r_{imk},\,\max_{m\in M_j}r_{jmk})$ is the maximum usage of resource $k\in K$ by either activity $i$ or $j$, and as before, $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan.
\end{comment}

The solution to he optimal objective value of the master problem (\ref{eqn:master_first})-(\ref{eqn:master_last}), denoted by $\eta^*$, forms a lower bound to the original robust MRCPSP problem. Note that for the first iteration, the solution to the master problem corresponds to an optimal solution to the MRCPSP with nominal activity durations.

@@

In line with \cite{bruni2017adjustable} and \cite{balouka2021robust}, two relaxations of the subproblem have been included in the master problem in order to strengthen it. The first relaxation is implemented through constraints (\ref{eqn:master_relaxation1_1}) and (\ref{eqn:master_relaxation1_2}). These simply enforce the extended precedence relationships using nominal activity durations for each processing mode. The second relaxation is captured by constraints (\ref{eqn:master_relaxation2_1})-(\ref{eqn:master_relaxation2_5}), which use flow variables $\phi_{ij}$ to form a longest-path problem through the extended project network using minimal activity durations.

@@

The subproblem at iteration $t$ considers the worst-case makespan for the mode choices and sufficient selection from the master problem at iteration $t$, denoted by $\bm{m}^{*t}$ and $X_{\bm{m}^{*t}}$ respectively. This is done by solving an uncertain longest-path problem through the extended project network, where activity durations lie in the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$. This problem can be formulated as 
\begin{align}
	V^{*t} = &\max \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \bar{d}_{im^{*t}_i}\alpha_{ij} + \hat{d}_{im^{*t}_i}w_{ij}\label{eqn:subproblem_first}\\
	& \sum_{(i,n+1)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{i,n+1} = 1\\
	& \sum_{(0,i)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{0,i} = 1\\
	& \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{ij} - \sum_{(j,i)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{ji} = 0 \qquad \forall i \in V\setminus\{0,n+1\}\\
	& w_{ij} \leq \xi_i \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& w_{ij} \leq \alpha_{ij} \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& \sum_{i\in V}\xi_i \leq \Gamma\\
	& \alpha_{ij}\in \{0,1\} \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& w_{ij}\geq 0 \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& 0\leq \xi_i \leq 1 \qquad \forall i\in V \label{eqn:subproblem_last}
\end{align}

Variables $\alpha_{ij}$ define the longest path through the network. This path, which we denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$ has length $V^{*t}$. Variables $\xi_i$ are the activity delay variables, whilst $w_{ij}$ are used to linearise the formulation.

@@

\begin{algorithm}[h] 
\caption{Benders' decomposition algorithm}
\begin{algorithmic}[1]
\Initialise{Set $LB=-\infty$, $UB=+\infty$ and $t=1$.}
\While{$UB > LB$}
	\State \textbf{Solve master problem (\ref{eqn:master_first})-(\ref{eqn:master_last})} 
	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m^{*t}}$ and precedences $\bm{y}_{\bm{m}^{*t}}$.
	\State \hspace{\algorithmicindent} If $\eta^{*t}>LB$, update $LB\leftarrow \eta^{*t}$ 
	\State \textbf{Solve subproblem (\ref{eqn:subproblem_first})-(\ref{eqn:subproblem_last})}
	\State \hspace{\algorithmicindent} Get objective value $V^{*t}$ and longest path $\pi^{*t}$
	\State \hspace{\algorithmicindent} If $V^{*t}<UB$, update $UB\leftarrow V^{*t}$
	\State \textbf{Add cut (\ref{eqn:benders_cut}) to master problem}
	\State Update $t\leftarrow t+1$
\EndWhile
\State \Return $UB$
\end{algorithmic}
\label{alg:benders}
\end{algorithm}

@@

Here we solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$ to illustrate the implementation of the Benders' decomposition approach which is summarised in Algorithm \ref{alg:benders}. This problem is solved in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}. Note that in the final iteration, the algorithm can terminate after solving the master problem and updating the $UB$ since $LB=UB$, i.e. there is no need to solve the subproblem in the final iteration.

>>>>

% bibliography and referencing
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{bibentry}  % inline referencing
\renewcommand{\inlinebibentry}[1]{\cite{#1} \bibentry{#1}}
\usepackage{hyperref}
\usepackage{xcolor}
\definecolor{darkgreen}{RGB}{0,127,173}
\definecolor{darkblue}{RGB}{0,0,200}
\hypersetup{colorlinks, linkcolor={darkblue}, citecolor={darkgreen}, urlcolor={darkblue}}

@@

\author[1]{Matthew Bold\footnote{Corresponding author, email: m.bold1@lancaster.ac.uk}}
\author[2]{Marc Goerigk}
\affil[1]{STOR-i Centre for Doctoral Training, Lancaster University, Lancaster, UK}	
\affil[2]{Network and Data Science Management, University of Siegen, Siegen, Germany}

@@

\nobibliography*  % tells bibentry to load bibliography database

\begin{abstract}
	This paper presents a mixed-integer linear programming formulation for the multi-mode resource-constrained project scheduling problem with uncertain activity durations. We consider a two-stage robust optimisation approach, where activity modes are selected and resource conflicts are resolved under the uncertain activity durations in a first-stage, which is followed by a second-stage where the schedule is evaluated once the activity durations become known. We aim to find solutions to minimise the worst-case project makespan, whilst assuming that activity durations lie in a budgeted uncertainty set.

	The proposed mixed-integer linear programming formulation is an extension of the formulation proposed by \cite{bold2021compact} (\bibentry{bold2021compact}) for the problem without multiple activity modes. Computational experiments show that this easy-to-implement formulation can be solved many times faster than the current state-of-the-art for solving this problem, solving over 40\% more instances to optimality over the same set of benchmarking instances. 
\end{abstract}

@@

We propose solving the robust MRCPSP using an extended version of the mixed-integer programming formulation developed by \cite{bold2021compact} for solving the robust RCPSP. This formulation combines the first and second-stage problems into a single compact formulation and was shown to significantly outperform the strongest iterative decomposition-based methods for the robust RCPSP.

This formulation is constructed by recasting the adversarial subproblem of determining the worst-case activity durations for a given set of mode choices and sufficient selection as a longest path problem on an augmented project network. This augmented project network is formed of $\Gamma+1$ connected copies of the original project network, with each level being used to account for a delay to a single activity in the project. Having reformulated the adversarial subproblem as a longest-path problem, it can be dualised an inserted into a formulation for the first-stage problem, to arrive at a complete compact formulation for the full problem. Given that the derivation of this formulation is more or less identical to the derivation of the corresponding formulation for the robust RCPSP, we omit it here and instead refer the reader to \cite{bold2021compact}.

In the resulting compact formulation, $y_{ij}$ variables are used to define a complete set of transitive precedences between the project activities that are formed by the original project precedence constraints and the additional precedences introduced to resolve resource conflicts, i.e. $\{(i,j),\, i,j\in V:y_{ij}=1\}=T(E\cup X_{\bm{m}})$, where $T(\cdot)$ is used to denote the transitive closure of a set. Variables $x_{im}$ determine the processing mode for each activity, whilst resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion. The $S_{i\gamma}$ variables are dual variables associated with the longest-path subproblem, which in this formulation these variables are used to track the start time of activities under the worst-case activity durations. Using these variables, the compact formulation for the robust MRCPSP can be written as   
\begin{align}
\min \ & S_{n+1,\Gamma}\label{eqn:compact1}\\
\text{s.t. } & S_{00}=0\label{eqn:compact2}\\
	     & S_{j\gamma} - S_{i\gamma} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i,\,\gamma=0,\dots,\Gamma\label{eqn:compact_bigM1}\\
	     & S_{j,\gamma+1} - S_{i\gamma} \geq (\bar{d}_{im} + \hat{d}_{im})x_{im} - N(1-y_{ij})\nonumber \\
	     & & \hspace{-8cm} \forall (i,j)\in V^2,\, m\in M_i,\,\gamma = 0,\dots,\Gamma-1\label{eqn:compact_bigM2}\\
	     & y_{ij}=1 \qquad \forall (i,j) \in E\cup\{(n+1,n+1)\}\label{eqn:compact5}\\
	     & y_{ij} + y_{ji} \leq 1 \qquad \forall i,j\in V,\, i<j\label{eqn:compact_trans1}\\
	     & y_{ij} = y_{ip} + y_{p j} - 1 \qquad \forall i,j,p\in V,\,i\neq j\neq p\label{eqn:compact_trans2}\\
	     & f_{ijk}\leq P_{ijk}y_{ij} \qquad \forall (i,j)\in V^2,\,i\neq n+1,\,j\neq0,\, \forall k \in K\label{eqn:compact6}\\
	     & \sum_{i\in V\setminus\{n+1\}}f_{ijk}=\sum_{m\in M_j}r_{jmk}x_{jm} \qquad \forall j \in V\setminus\{0\} ,\, \forall k\in K\label{eqn:compact7}\\
	     & \sum_{j\in V\setminus\{0\}}f_{ijk}=\sum_{m\in M_i}r_{imk}x_{im} \qquad \forall i \in V\setminus\{n+1\},\,\forall k\in K\label{eqn:compact8}\\
	     & \sum_{m\in M_i}x_{im} = 1 \qquad \forall i\in V\label{eqn:compact9}\\
	     & \sum_{m\in M_i}r^{'}_{imk}x_{im} \leq R^{'}_k \qquad \forall i \in V,\,k\in K^{'}\label{eqn:compact_nonrenewable}\\
	     & S_{i\gamma}\geq 0 \qquad \forall i\in V,\,\gamma \in 0,\dots, \Gamma\label{eqn:compact10}\\
	     & y_{ij}\in\{0,1\} \qquad \forall (i,j)\in V^2\label{eqn:compact11}\\
	     & f_{ijk}\geq 0 \qquad \forall (i,j)\in V^2 ,\, \forall k \in K\label{eqn:compact12}\\
	     & x_{im}\in\{0,1\} \qquad \forall i\in V,\,m\in M_i\label{eqn:compact13},
\end{align}
where $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan, and $P_{ijk}=\min\{\max_{m\in M_i}r_{imk},\,\max_{m\in M_j} r_{jmk}\}$ is the maximum possible flow of resource $k$ from $i$ into $j$.

@@

With polynomially many variables and constraints this formulation can be implemented straightforwardly using standard mathematical optimisation software such as Gurobi or CPLEX.

@@

In this section we outline an alternative approach to solving the robust MRCPSP based on Benders' decomposition. Such a solution method for the robust MRCPSP was first presented in \cite{balouka2021robust}, and as mentioned previously, this itself was based on the approach to solve the robust RCPSP used in \cite{bruni2017adjustable}.

The main idea behind this approach is to decompose the full problem into a two smaller problems: 1. a master problem that determines activity processing modes and resolves resource conflicts, and 2. a subproblem that takes the solution from the master problem and evaluates it by finding the worst-case makespan for the resulting network by solving a longest path problem. The solution to the master problem forms a lower bound to the optimal objective value of the original problem, whilst the solution to the subproblem provides an upper bound. These two problems are solved iteratively, with optimality cuts being added to the master problem each iteration based on the solution to the subproblem until the lower and upper bounds meet.

After initially implementing the Benders' algorithm as it is presented in \cite{balouka2021robust}, it was discovered that its performance could be strengthened by replacing the master problem formulation used in that implementation with a stripped-down version of the compact formulation (\ref{eqn:compact1})-(\ref{eqn:compact13}) in which the uncertainty is removed. In this section we therefore present our strengthened implementation of the Benders' approach for solving the robust MRCPSP. Note that a comparison of these two Benders' implementations is included in Section \ref{section:experiments}. For the details of the original implementation of the Benders' decomposition algorithm for the robust MRCPSP, see \cite{balouka2021robust}.

@@

The role of the master problem is to determine a choice of activity processing modes $\bm{m}$, as well as a sufficient selection $X_{\bm{m}}$ to resolve any resulting resource conflicts. This problem is solved without the direct consideration of the uncertain activity durations, and instead, uncertainty is accounted for in the subproblem and communicated back to the master problem with the use of optimality cuts. Hence, to remove the uncertainty from the model, (\ref{eqn:compact1})-(\ref{eqn:compact13}) is modified by considering only a single level of the augmented project network. Using this formulation, the master problem can be written as
\begin{align}
\min \ & \eta\label{eqn:master_first}\\
\text{s.t. } & \eta \geq S_{n+1}\\
	     & S_{0}=0\label{eqn:master2}\\
	     & S_{j} - S_{i} \geq \bar{d}_{im}x_{im} - N(1-y_{ij}) \qquad \forall (i,j)\in V^2,\,\forall m\in M_i\label{eqn:master3}\\
	     & \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=1,\dots,t-1\label{eqn:master_cuts}\\
	     & \textnormal{(\ref{eqn:compact5})-(\ref{eqn:compact_nonrenewable}),\,(\ref{eqn:compact11})-(\ref{eqn:compact13})}\\
	     & S_i\geq 0 \qquad \forall i\in V,\label{eqn:master_last}
\end{align}
where (\ref{eqn:master_cuts}) are the optimality cuts generated from the solutions of the resulting subproblems at each of the previous iterations.

The optimal objective value of the master problem (\ref{eqn:master_first})-(\ref{eqn:master_last}), denoted by $\eta^*$, forms a lower bound to the original robust MRCPSP problem. Note that for the first iteration, the solution to the master problem corresponds to an optimal solution to the MRCPSP with nominal activity durations.

% Master problem formulation from Balouka & Cohen (2021)
\begin{comment}
The processing mode for each activity is determined by binary variables $x_{im}$. The $y_{ij}$ variables define a complete set of precedences between the project activities, either due to the original project precedence constraints, or due to the resolution of resource conflicts, i.e. $\{(i,j),\, i,j\in V:y_{ij}=0\}=T(E\cup X_{\bm{m}})$, where $T(E\cup X_{\bm{m}})$ is the transitive closure of the extended project precedence network. Resource flow variables $f_{ijk}$ track the amount of renewable resource $k$ that is transferred from activity $i$ to activity $j$ upon its completion. Finally, $s_i$ and $\phi_{ij}$ are additional variables used to define two relaxations of the subproblem to strengthen the master problem formulation. Using these variables, the master problem can be formulated as follows:

\begin{align}
	& \min \eta\label{eqn:master_first}\\
	& y_{ij} = 1 \qquad \forall (i,j)\in E \label{eqn:master_y1}\\
	& y_{ij} + y_{ji} \leq 1 \qquad \forall i,j\in V,\, i<j\label{eqn:master_y2}\\
	& y_{ij} = y_{ip} + y_{p j} - 1 \qquad \forall i,j,p\in V,\,i\neq j\neq p\label{eqn:master_y3}\\
	& f_{ijk} \leq \min(r_{im_ik}, r_{jm_jk})\cdot y_{ij} + (1 - x_{i,m_i})(\bar{r}_{ijk} - \min(r_{im_ik}, r_{jm_jk}))\nonumber\\
	& & \hspace{-8cm} + (1 - x_{j,m_j})(\bar{r}_{ijk} - \min(r_{im_ik}, r_{jm_jk})) \nonumber \\
	& & \hspace{-10cm}\forall i\in V\setminus\{n+1\},\, j \in V\setminus\{0\},\, i\neq j,\,m_i\in M_i,\,m_j\in M_j,\,k\in K\label{eqn:master_f1}\\
	& \sum_{i\in V\setminus\{j,n+1\}}f_{ijk} = \sum_{m\in M_j}r_{jmk}x_{jm} \qquad \forall j \in V\setminus\{0\} ,\,k\in K\label{eqn:master_f2}\\
	& \sum_{j\in V\setminus\{0,i\}}f_{ijk} = \sum_{m\in M_i}r_{imk}x_{im} \qquad \forall i \in V\setminus\{n+1\} ,\,k\in K\label{eqn:master_f3}\\
	& \sum_{m\in M_i}x_{im} = 1\qquad \forall i \in V\label{eqn:master_x}\\
	& \sum_{m\in M_i}r^{\nu}_{imk}x_{im} \leq R^{\nu}_k \qquad \forall i\in V,\,k\in K^{\nu}\label{eqn:master_nonrenewable}\\
	& \eta \geq \lambda(x, y, M^{*\ell}) \qquad \forall \ell=0,\dots,t-1\label{eqn:master_cuts}\\
	& s_j - s_i \geq \bar{d}_{im} - N(1-y_{ij}) - N(1-x_{im}) \qquad \forall i,j\in V,\, i\neq j,\,m\in M_i\label{eqn:master_relaxation1_1}\\
	& \eta \geq s_{n+1}\label{eqn:master_relaxation1_2}\\
	& \eta \geq \sum_{i \in V}\sum_{j\in V}\Big(\min_{m\in M_i}\bar{d}_{im}\Big)\phi_{ij}\label{eqn:master_relaxation2_1}\\
	& \sum_{j\in V}\phi_{ji} = \sum_{j\in V}\phi_{ij} \qquad \forall i \in V\setminus\{0,n+1\}\\
	& \sum_{j\in V}\phi_{0j} = 1\\
	& \sum_{j\in V}\phi_{j,n+1} = 1\\
	& \phi_{ij} \leq y_{ij} \qquad \forall i,j\in V,\,i\neq j\label{eqn:master_relaxation2_5}\\
	& y_{ij} \in\{0,1\} \qquad \forall i,j\in V,\,i\neq j\\
	& x_{im} \in \{0,1\} \qquad \forall i \in V,\, m\in M_i\\
	& f_{ijk} \geq 0 \qquad \forall i\in V\setminus\{n+1\},\,j\in V\setminus\{0\},\,i\neq j,\,k\in K\\
	& s_i \geq 0 \qquad \forall i\in V\\
	& \phi_{ij} \geq 0 \qquad \forall i,j\in V \label{eqn:master_last}
\end{align}
where $\bar{r}_{ijk}=\max(\max_{m\in M_i}r_{imk},\,\max_{m\in M_j}r_{jmk})$ is the maximum usage of resource $k\in K$ by either activity $i$ or $j$, and as before, $N=\sum_{i\in V}\max_{m\in M_i}(\bar{d}_{im}+\hat{d}_{im})$ is an upper bound on the minimum makespan.

@@

In line with \cite{bruni2017adjustable} and \cite{balouka2021robust}, two relaxations of the subproblem have been included in the master problem in order to strengthen it. The first relaxation is implemented through constraints (\ref{eqn:master_relaxation1_1}) and (\ref{eqn:master_relaxation1_2}). These simply enforce the extended precedence relationships using nominal activity durations for each processing mode. The second relaxation is captured by constraints (\ref{eqn:master_relaxation2_1})-(\ref{eqn:master_relaxation2_5}), which use flow variables $\phi_{ij}$ to form a longest-path problem through the extended project network using minimal activity durations.
\end{comment}

@@

The subproblem at iteration $t$ considers the worst-case makespan for the mode choices and sufficient selection from the master problem at iteration $t$, denoted by $\bm{m}^{*t}$ and $X_{\bm{m}^{*t}}$ respectively. This is done by solving an uncertain longest-path problem through the extended project network $T(E\cup X_{\bm{m}^{*t}})$, where activity durations lie in the budgeted uncertainty set $\U_{\bm{m}^{*t}}(\Gamma)$, which can be formulated as 
\begin{align}
	V^{*t} = &\max \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \bar{d}_{im^{*t}_i}\alpha_{ij} + \hat{d}_{im^{*t}_i}w_{ij}\label{eqn:subproblem_first}\\
	& \sum_{(i,n+1)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{i,n+1} = 1\\
	& \sum_{(0,i)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{0,i} = 1\\
	& \sum_{(i,j)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{ij} - \sum_{(j,i)\in T(E\cup X_{\bm{m}^{*t}})} \alpha_{ji} = 0 \qquad \forall i \in V\setminus\{0,n+1\}\\
	& w_{ij} \leq \xi_i \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& w_{ij} \leq \alpha_{ij} \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& \sum_{i\in V}\xi_i \leq \Gamma\\
	& \alpha_{ij}\in \{0,1\} \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& w_{ij}\geq 0 \qquad \forall (i,j)\in T(E\cup X_{\bm{m}^{*t}})\\
	& 0\leq \xi_i \leq 1 \qquad \forall i\in V. \label{eqn:subproblem_last}
\end{align}

Variables $\alpha_{ij}$ define the longest path through the network. This path, which is denote as $\pi^{*t}=\{(i,j),\,i,j\in V:\alpha_{ij}=1\}$ has length $V^{*t}$. Variables $\xi_i$ are the activity delay variables, whilst $w_{ij}$ are used to linearise the formulation. Note that this formulation of the subproblem is identical to the one implemented by \cite{balouka2021robust}.

@@

@@

Here, to demonstrate its implementation, we use the Benders' decomposition approach to solve the example shown in Figure \ref{fig:mrcpsp_network} with $R_1=4$ and $\Gamma=2$. An outline of the implementation of the Benders' approach is presented in Algorithm \ref{alg:benders}.

The algorithm solves this instance in 6 iterations, and the solution information from the master and subproblem at each of these iterations is shown in Table \ref{table:benders_example}. Note that in the final iteration, the algorithm can terminate after solving the master problem and updating the $UB$ since $LB=UB$, i.e. there is no need to solve the subproblem in the final iteration.

\begin{algorithm}[h] 
\caption{Benders' decomposition algorithm.}
\begin{algorithmic}[1]
\Initialise{Set $LB=-\infty$, $UB=+\infty$ and $t=1$.}
\While{$UB > LB$}
	\State \textbf{Solve master problem (\ref{eqn:master_first})-(\ref{eqn:master_last})} 
	\State \hspace{\algorithmicindent} Get objective value $\eta^{*t}$, processing modes $\bm{m^{*t}}$ and precedences $\bm{y}_{\bm{m}^{*t}}$
	\State \hspace{\algorithmicindent} If $\eta^{*t}>LB$, update $LB\leftarrow \eta^{*t}$ 
	\State \textbf{Solve subproblem (\ref{eqn:subproblem_first})-(\ref{eqn:subproblem_last})}
	\State \hspace{\algorithmicindent} Get objective value $V^{*t}$ and longest path $\pi^{*t}$
	\State \hspace{\algorithmicindent} If $V^{*t}<UB$, update $UB\leftarrow V^{*t}$
	\State \textbf{Add cut (\ref{eqn:benders_cut}) to master problem}
	\State Update $t\leftarrow t+1$
\EndWhile
\State \Return $UB$
\end{algorithmic}
\label{alg:benders}
\end{algorithm}
